---
title: "Aides Drug"
author: "Andrea Calderon"
date: "Feb 26, 2024"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

$H_0$: Those who have been diagnosed with aids do not appear to benefit from the new preventative medication.


$H_A$: Aids patients appear to respond well to the recently developed prophylactic medication.



# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->


Variable we use are:  

treatment - categorical   
DrugUser - categorical   
infected - categorical  

Infected depends on treament for DrugUsers. 

We get just the DrugUsers from the population. 


```{r}
drugUsersData= select(triUmp, treatment, DrugUser, infected)
drugUsersData= filter(drugUsersData,DrugUser=='yes')
```




# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->


```{r}
barchartGC(~infected + treatment,data=drugUsersData)
barchartGC(~infected + treatment,data=drugUsersData, type="percent")

```

 By looking at the chart you are able to see that the charts shown you can see that the placebo works.The percentage has risen where as those who have been getting the drug shows a lower percentage showing that the drug is working.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=drugUsersData)
rowPerc(table1)
colPerc(table1)
```
When looking into this graph you are able to see if you are infected with aids then 24.29% were getting the drug and 75.71% were getting the placebo. 


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

``` {r}
chisq.test (table1)
```
```{r} 
chisqtestGC(table1)
```

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
